Usability of Mobile Map Applications

### 2014-10-13 usability-study/usability-study.tex

\vspace{-.2in}
\begin{center}
\begin{abstract}
\noindent This investigation is motivated by the increasing significance of the usability of systems in today's competitive technology market. Usability testing and usability evaluations have become important tools in the assessment of user interfaces. This paper presents empirical mobile usability studies of two distinct map applications: Google Maps and Apple Maps. Participants of this investigation were asked to carry out three tasks with each mobile map application. The results of this examination include (a) the contextual factors studied; (b) quantitative data collected from each task; (c) a brief qualitative review of the applications from each participant; and (d) the core usability metrics measured. This paper not only presents the usability metrics collected by the study but also presents a heuristic evaluation of each map application.
\end{abstract}
\end{center}

@@

\section{Introduction}
When the Apple Maps mobile application was first released as part of iOS 6 in September 2012, it received a significant amount of negative feedback. Walt Mossberg of the \textit{Wall Street Journal} was one of many who thought that Apple Maps was the biggest drawback of the iPhone 5.\cite{Mossberg} At this time, the Google Maps mobile apps, on both iOS and Android, were used by roughly 81 million people.\cite{ComScore} However, Apple Maps grew, matured, and began gaining traction. By September 2013, Apple Maps gained 35 million regular users. The number of users of Google Maps dropped to around 58.7 million at this time.\cite{ComScore} Although the Apple Maps mobile app is newer than the Google Maps mobile app,\footnote{The initial release of Google Maps for Mobile was in September of 2008.} the two map applications are comparable with respect to features they provide as well as popularity nowadays.
\medskip
\par
Since both Google Maps apps and Apple Maps apps are widely known and used, it is important that the usability metrics of the two systems are examined. Thus, this study aims to empirically investigate the usability of each application. The current consensus within the field of Interaction Design is that usability depends on five distinct metrics: \textit{learnability}, \textit{efficiency}, \textit{errors}, \textit{memorability}, and \textit{satisfaction}.\cite{Nielsen} The three metrics on which this particular study focuses are efficiency, errors, and satisfaction.
\section{Contextual Factors}
Prior to having each test subject perform the tasks, a little contextual data about the subject was first collected. Users were asked about their previous experience with Apple Maps as well as their previous experience with Google Maps.
\vspace{-.2in}
\begin{figure}[ht]
\begin{center}
\includegraphics[keepaspectratio, width=.7\textwidth ]{user_context.pdf}
\end{center}
\end{figure}
\clearpage
\par
The fact that nearly every test subject had significantly more Google Maps experience than Apple Maps experience was a challenge in the conducting of this study.  Since this study aimed to test the usability metric of efficiency, and \textit{not} that of learnability, we did not test subjects if they had no prior experience with the application. For future studies, however, it might be worthwhile to give inexperienced users a short tutorial so it is possible to test them thereafter.

\section{The Experiment} \label{sec:Experiment}
Each person who participated in this study had to carry out three concrete tasks with the map application[s] with which they were experienced.
\begin{enumerate}
  \item Save a pin anywhere on the map and get driving directions to that pin.
  \item Find the location of the nearest Vons.
  \item Find the fastest route from the current location to LACMA.
\end{enumerate}
In order to keep the study as controlled as possible, all testing was done on iOS devices (no map applications of android devices were tested). As test subjects executed the tasks, information was collected pertaining to their efficiency, errors, and satisfaction.

\vspace{-.1in}
\section{Usability Metrics}
\vspace{-.1in}
\subsection{Efficiency}
\par
\textbf{First Task: }Drop a pin on the map and get driving directions to that pin.
\vspace{-.4in}
\begin{figure}[ht]
\begin{center}
\vspace{-.1in}
\includegraphics[keepaspectratio, width=.8\textwidth ]{task1.pdf}
\end{center}
\end{figure}
\begin{center}
\vspace{-.6in}
\par
Average time for task \#1 on Google Maps app: $ \approx $ \textit{21.08 seconds}
\par
Average time for task \#1 on Apple Maps app: $ \approx $ \textit{19.24 seconds}
\end{center}
\par
\noindent
This task's results suggest that on average, Apple Maps is more efficient for dropping pins and mapping to them.\footnote{Though subjects using Apple Maps were more efficient \textit{on average}, it notable that there were a couple of anomalies within the Google Maps users who significantly raised the average.}
\clearpage
\par
\noindent
\textbf{Second Task: }Find the location of the nearest Vons.
\begin{figure}[htbp]
\begin{center}
\vspace{-.4in}
\includegraphics[keepaspectratio, width=.8\textwidth ]{task2.pdf}
\end{center}
\end{figure}
\begin{center}
\vspace{-.6in}
\par
Average time for task \#2 on Google Maps app: $ \approx $ \textit{20.53 seconds}
\par
Average time for task \#2 on Apple Maps app: $ \approx $ \textit{23.62 seconds}
\end{center}
\par
\noindent
This task's results suggest that on average, Google Maps is more efficient for finding the desired locations nearest to your current location. 
\par
\medskip
\noindent
\textbf{Third Task: }Find the fastest route from the current location to LACMA.
\begin{figure}[htbp]
\begin{center}
\vspace{-.4in}
\includegraphics[keepaspectratio, width=.8\textwidth ]{task3.pdf}
\end{center}
\end{figure}
\begin{center}
\vspace{-.6in}

@@

\end{table}
\par
The results suggest that errors are less likely with Google Maps .

@@

The icons of the Google Maps mobile application comply with the Google Design Guidelines\cite{Google}: simple, modern, and geometric in execution. Thus, Google Maps chooses to not use the standard iOS icons\cite{Apple}:

@@

Though it is not required to use these provided iOS icons, the iOS HIG (Human Interface Guidelines) state that it is ``a good idea to use the built-in icons as much as possible because users already know what they mean."\cite{Apple} Balancing consistency in platform with consistency of brand is a challenge, but there is a possibility that adhering more to the iOS HIG could improve usability of Google Maps even more. In this study, there was a higher percentage of errors specifically involving clicking wrong icons in the Google Maps application, which could have possibly been prevented if standard iOS icons were used. 

@@

This suggests that the icon on the top left-hand corner ---which happens to be the directions screen in the Apple Maps app--- is the most important. I would have to agree that getting directions is a very important functionality. Thus, it seems as if the app is complying to the guidelines. However, a couple of test subjects tried to click on the icon second from the right on the bottom for directions. It is possible that this guideline of important functionality being `elevated' may not be intuitive to all users.

@@

The Apple Maps mobile application certainly complies to the guideline of placing important functionality toward the upper-left corner of the screen. Although this is a guideline provided by Apple, it is not entirely clear whether or not this is the best placement for important functionality. The top corner opposite to the hand the user holds the phone with requires a little extra effort and stretching for many hands. This adds an additional variable in an equation known as Fitts' Law\cite{Fitts}, which is indicative of an increase in the movement time. The motion between two targets is no longer a fluid, resistance-free motion. It is possible that, even though it complies with Apple guidelines, the placement of the directions button caused errors that could be prevented by a different placement. In a future study, it would be noteworthy to deduce what placement of important functionality is optimal for efficiency, error prevention, and satisfaction.

@@

>>>>

\vspace{-.2in}
\begin{center}
\begin{abstract}
\noindent This investigation is motivated by the increasing significance of the usability of systems in today's competitive technology market. Usability testing and evaluation have become important tools in the assessment of user interfaces. This paper presents empirical mobile usability studies of two distinct map applications: Google Maps and Apple Maps. Participants of this investigation were asked to carry out three tasks with each mobile map application. The results of this examination include (a) the contextual factors studied; (b) quantitative data collected from each task; (c) a brief qualitative review of the applications from each participant; and (d) the core usability metrics measured. This paper not only presents the usability metrics collected by the study but also presents a heuristic evaluation of each map application.
\end{abstract}
\end{center}

@@

\section{Introduction}
When the Apple Maps mobile application was first released as part of iOS 6 in September 2012, it received a significant amount of negative feedback. Walt Mossberg of the \textit{Wall Street Journal} was one of many who thought that Apple Maps was the biggest drawback of the iPhone 5 \cite{Mossberg}. At this time, the Google Maps mobile apps, on both iOS and Android, were used by roughly 81 million people.\cite{ComScore}. However, Apple Maps grew, matured, and began gaining traction. By September 2013, Apple Maps gained 35 million regular users. The number of users of Google Maps dropped to around 58.7 million at this time \cite{ComScore}. Although the Apple Maps mobile app is newer than the Google Maps mobile app,\footnote{The initial release of Google Maps for Mobile was in September of 2008.} the two map applications are comparable with respect to features they provide as well as popularity nowadays.
% JD: Excellent contextualization of the study with objective data to back it up.

\medskip % JD: Shouldn't be necessary.  If LaTeX doesn't do it this way, there is likely
\par     %     a package that changes its behavior, rather than inserting one-offs like this.
Since both Google Maps apps and Apple Maps apps are widely known and used, it is important that the usability metrics of the two systems are examined. Thus, this study aims to empirically investigate the usability of each application. The current consensus within the field of Interaction Design is that usability depends on five distinct metrics: \textit{learnability}, \textit{efficiency}, \textit{errors}, \textit{memorability}, and \textit{satisfaction} \cite{Nielsen}. The three metrics on which this particular study focuses are efficiency, errors, and satisfaction.
% JD: I won't change anymore at this point, but citations typically appear *before* the period.

\section{Contextual Factors}
Prior to having each test subject perform the tasks, a little contextual data about the subject was first collected. Users were asked about their previous experience with Apple Maps as well as their previous experience with Google Maps.
\vspace{-.2in}
\begin{figure}[ht]
\begin{center}
\includegraphics[keepaspectratio, width=.7\textwidth ]{user_context.pdf}
\end{center}
\end{figure}
\clearpage
\par
The fact that nearly every test subject had significantly more Google Maps experience than Apple Maps experience was a challenge in the conducting of this study.  Since this study aimed to test the usability metric of efficiency, and \textit{not} that of learnability, we did not test subjects if they had no prior experience with the application. For future studies, however, it might be worthwhile to give inexperienced users a short tutorial so it is possible to test them thereafter.

\section{The Experiment} \label{sec:Experiment}
Each person who participated in this study had to carry out three concrete tasks with the map application[s] with which they were experienced.
\begin{enumerate}
  \item Save a pin anywhere on the map and get driving directions to that pin.
  \item Find the location of the nearest Vons.
  \item Find the fastest route from the current location to LACMA.
\end{enumerate}
In order to keep the study as controlled as possible, all testing was done on iOS devices (no map applications of Android devices were tested). As test subjects executed the tasks, information was collected pertaining to their efficiency, errors, and satisfaction.

% JD: Yes...shouldn't be necessary to insert this.  LaTeX's intent is to relieve you
%     of this concern.  If you don't like what you see, leave it be unless it materially
%     interferes with the communication of your content.
\vspace{-.1in}
\section{Usability Metrics}
\vspace{-.1in}
\subsection{Efficiency}
\par
\textbf{First Task: }Drop a pin on the map and get driving directions to that pin.
\vspace{-.4in}
\begin{figure}[ht]
\begin{center}
\vspace{-.1in}
\includegraphics[keepaspectratio, width=.8\textwidth ]{task1.pdf}
\end{center}
\end{figure}
\begin{center}
\vspace{-.6in}
\par
Average time for task \#1 on Google Maps app: $ \approx $ \textit{21.08 seconds}
\par
Average time for task \#1 on Apple Maps app: $ \approx $ \textit{19.24 seconds}
\end{center}
\par
\noindent
This task's results suggest that on average, Apple Maps is more efficient for dropping pins and mapping to them.\footnote{Though subjects using Apple Maps were more efficient \textit{on average}, it notable that there were a couple of anomalies within the Google Maps users who significantly raised the average.} % JD: For more expansive studies, statistical methods help clean this up.
\clearpage
\par
\noindent
\textbf{Second Task: }Find the location of the nearest Vons.
\begin{figure}[htbp]
\begin{center}
\vspace{-.4in}
\includegraphics[keepaspectratio, width=.8\textwidth ]{task2.pdf}
\end{center}
\end{figure}
\begin{center}
\vspace{-.6in}
\par
Average time for task \#2 on Google Maps app: $ \approx $ \textit{20.53 seconds}
\par
Average time for task \#2 on Apple Maps app: $ \approx $ \textit{23.62 seconds}
\end{center}
\par
\noindent
This task's results suggest that on average, Google Maps is more efficient for finding the desired locations nearest to your current location. 
\par
\medskip
\noindent
\textbf{Third Task: }Find the fastest route from the current location to LACMA.
\begin{figure}[htbp]
\begin{center}
\vspace{-.4in}
\includegraphics[keepaspectratio, width=.8\textwidth ]{task3.pdf}
\end{center}
\end{figure}
\begin{center}
\vspace{-.6in}

@@

\end{table}
\par
The results suggest that errors are less likely with Google Maps.

@@

The icons of the Google Maps mobile application comply with the Google Design Guidelines \cite{Google}: simple, modern, and geometric in execution. Thus, Google Maps chooses to not use the standard iOS icons \cite{Apple}:

@@

Though it is not required to use these provided iOS icons, the iOS HIG (Human Interface Guidelines) state that it is ``a good idea to use the built-in icons as much as possible because users already know what they mean \cite{Apple}.'' Balancing consistency in platform with consistency of brand is a challenge, but there is a possibility that adhering more to the iOS HIG could improve usability of Google Maps even more. In this study, there was a higher percentage of errors specifically involving clicking wrong icons in the Google Maps application, which could have possibly been prevented if standard iOS icons were used. % JD: Nice, well-connected conclusion.

@@

This suggests that the icon on the top left-hand corner ---which happens to be the directions screen in the Apple Maps app--- is the most important. I would have to agree that getting directions is a very important functionality. Thus, it seems as if the app is complying with the guidelines. However, a couple of test subjects tried to click on the icon second from the right on the bottom for directions. It is possible that this guideline of important functionality being `elevated' may not be intuitive to all users.

@@

The Apple Maps mobile application certainly complies to the guideline of placing important functionality toward the upper-left corner of the screen. Although this is a guideline provided by Apple, it is not entirely clear whether or not this is the best placement for important functionality. The top corner opposite to the hand the user holds the phone with requires a little extra effort and stretching for many hands. This adds an additional variable in an equation known as Fitts' Law \cite{Fitts}, which is indicative of an increase in the movement time. The motion between two targets is no longer a fluid, resistance-free motion. It is possible that, even though it complies with Apple guidelines, the placement of the directions button caused errors that could be prevented by a different placement. In a future study, it would be noteworthy to deduce what placement of important functionality is optimal for efficiency, error prevention, and satisfaction.
% JD: Yes, I think this notion of "elevation" needs to be examined alongside the
%     assorted "navigation models" that iOS describes (hierarchical, tabs, etc.).
%     There may be some interactions between these two that determine how an element
%     needs to be placed.
%
%     Meanwhile, nice application of Fitts' Law :)

@@

% JD: Not a bad start; next level would be using BibTeX.
